A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds

### 2013-06-10 paper/paper.tex
more writing.

\begin{lem}[Trout] If $v$ and $w$ are vertices in a positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
\end{lem}

@@

\begin{thm}[Trout] For any positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
\label{thm:discrete_BM_expanded_paths}
\end{thm}

@@

\noindent The proof of this theorem relies on the remarkable work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds along with a list of their topological types, see \cite{Lutz_online_manifolds} and \cite{Lutz_online_topological_types}. This census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$. Note that all manifolds are either homeomorphic to a 3-sphere or to the quotient space of a 3-sphere under the action of a finite group.

Here is a table of the number of triangulations in the census at each topological type and diameter.

@@

\noindent Compare this table with the corresponding counts indexed by the coarser {\em edge-diameter}.

\begin{figure}
\centering
\begin{tabular} {| r | r | r | r | r | r | r |}
\hline
Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
\hline
\hline
$1$ &  &  &  &   &3    & 3    \\
$2$ & 1&  &1 &17 &574  & 593  \\
$3$ &  &1 &1 &5  &3498 & 3505 \\
$4$ &  &  &  &   &675  & 675   \\
$5$ &  &  &  &   &11   & 11   \\
\hline
Total&1 &1 &2 &22   &4761   &4787  \\
\hline
\end{tabular}
\caption{The positively curved combinatorial 3-manifolds by edge-diameter and topological type}
\label{fig:manifold_counts}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
    \caption{Global statistics. TO DO: CHANGE DIAMETER LABELS TO $1, H, 2, J, 3, 2H, 4, 2J$ AND PUT THESE ON HORIZONTAL AXIS. TO DO: CHANGE GRAPH TO SHOW SEPARATE DATA POINTS FOR EACH TOPOLOGICAL TYPE. TO DO: CREATE ANOTHER GRAPH WITH JUST EDGE DIAMETER COUNTS, WHERE THE CORRESPONDENCE BETWEEN DIAMETER(s) AND EDGE-DIAMETER IS $0\rightarrow 0, 1\rightarrow 1, \{H, 2\}\rightarrow 2, \{J,3\}\rightarrow 3, 4\rightarrow 4, 2J\rightarrow 5$}
    \end{center}
    \label{global_statistics}
\end{figure}

>>>>

\begin{lem}[Trout] If $v$ and $w$ are vertices in a connected positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
\end{lem}

@@

\begin{thm}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
\label{thm:discrete_BM_expanded_paths}
\end{thm}

@@

\noindent The proof of this theorem relies on the remarkable work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds along with a list of their topological types, see \cite{Lutz_online_manifolds} and \cite{Lutz_online_topological_types}. This census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$. Note that all of these manifolds are homeomorphic to either a 3-sphere or to the quotient space of a 3-sphere under the action of a finite group.

A small Python program was created to examine each manifold in the census and compute its diameter. See Figure \ref{fig:global_statistics} for a plot of the number of manifolds by topological type and diameter. Note that, as claimed, every manifold with diameter greater than $J$ is a 3-sphere. The corresponding data using the coarser-grained edge-diameter can be found in Figure \ref{fig:global_statistics_edges}.

@@

\begin{figure}
\centering
\begin{tabular} {| r | r | r | r | r | r | r |}	
\hline
Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
\hline
\hline
$1$ &  &  &  &   &3    & 3    \\
$2$ & 1&  &1 &17 &574  & 593  \\
$3$ &  &1 &1 &5  &3498 & 3505 \\
$4$ &  &  &  &   &675  & 675   \\
$5$ &  &  &  &   &11   & 11   \\
\hline
Total&1 &1 &2 &22   &4761   &4787  \\
\hline
\end{tabular}
\caption{The positively curved combinatorial 3-manifolds by edge-diameter and topological type}
\label{fig:manifold_counts_edges}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
    \caption{Global statistics. TO DO: CHANGE DIAMETER LABELS TO $1, H, 2, J, 3, 2H, 4, 2J$ AND PUT THESE ON HORIZONTAL AXIS. TO DO: CHANGE GRAPH TO SHOW SEPARATE DATA POINTS FOR EACH TOPOLOGICAL TYPE. TO DO: CREATE ANOTHER GRAPH WITH JUST EDGE DIAMETER COUNTS}
    \end{center}
    \label{fig:global_statistics}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
    \caption{...TO DO}
    \end{center}
    \label{fig:global_statistics_edges}
\end{figure}
